My title

### 2024-03-11 paper/paper.qmd
add data section, add r package to bib

library(tidyverse)
library(palmerpenguins)
library(here)
library(rstanarm)
library(modelsummary)
library(ggplot2)
library(knitr)
library(marginaleffects)
library(plotly)
library(tibble)
library(margins)

@@

This report consists of four sections, not including the introduction. In Section 2, we look at the data used for our report and include some tables and graphs to illustrate the different groups of people in our data. In section 3, we build the model and discuss its justification and explanation. The results of our predictions is highlighted in the Section 4 using tables and graphs. Lastly, discussions are conducted based on the findings, which addresses the voting prediction results based on race, region, and employment status.

In this report, R statistical programming language is used, with R packages: [a lot of citations needed]

@@

@@

## Summary Statistics

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
# read in data
ces2020 <-  read_csv(here::here("data/analysis_data/cleaned_ces2020.csv"))
```

```{r}
# change column type to factor
ces2020 <-
  ces2020 |>
  mutate(
    voted_for = as_factor(voted_for),
    race = factor(
      race,
      levels = c(
        "White",
        "Black",
        "Hispanic",
        "Asian",
        "Native American",
        "Middle Eastern",
        "Two or more races",
        "Other"
      )
    ),
    region = factor(
      region,
      levels = c(
        "Northeast",
        "Midwest",
        "South",
        "West"
      )
    ),
    employ = factor(
      employ,
      levels = c(
        "Full-time",
        "Part-time",
        "Temporarily laid off",
        "Unemployed",
        "Retired",
        "Permanently disabled",
        "Homemaker",
        "Student",
        "Other"
      )
    )
  ) |>
  select(voted_for, race, region, employ)
```

# Model

@@

-------------------------------needs some explanation
```{r}
# We could use posterior predictive checks, introduced in Section 12.4, to show that the logistic regression approach is a good choice for this circumstance.
political_preferences1 <-
  readRDS(file = here::here("models/political_preferences1.rds"))

@@

# Results

@@

# # generate predictions for each of our grouped cells
# props$estimate <- predict.glm(political_preferences1, newdata = props,
#                           type = "response")
# 
# # create a new column with number of voters from each cell voting Biden
# props <- props %>% mutate(num_voters = n*estimate)
# 
# # get errors for our prediction
# errors <- predict.glm(model, newdata = props,
#                       type = "response", se.fit = T)
# lower <- errors$fit - errors$se.fit
# upper <- errors$fit + errors$se.fit
# 
# 
# # combine the proportions and the errors we generated
# props_error <- cbind(props, lower, upper)
# props_error <- props_error %>% 
#   rename("lower" = ...11,
#          "upper" = ...12)
# 
# # create a new column with number of voters from each cell voting Biden
# props_error <- props_error %>% mutate(num_voters_lower = n*lower,
#                                 num_voters_upper = n*upper)
# 
# # create a new dataset which contains upper and lower estimates for proportions
# # of votes for each state
# voting_biden <- props_error %>% mutate(biden_predict_prop=estimate*prop,
#                                        biden_predict_prop_lower = lower*prop,
#                                        biden_predict_prop_upper = upper*prop) %>%
#   group_by(stateicp) %>%
#   summarise(biden_predict = sum(biden_predict_prop),
#             biden_predict_lower = sum(biden_predict_prop_lower),
#             biden_predict_upper = sum(biden_predict_prop_upper))
```

```{r stateplot, echo=F, message=F, warning=F, fig.cap="Proportion of each State Voting for Biden"}
# # get state by state proportions for our polling data
# polling_props <- polling %>% 
#   group_by(stateicp, vote_biden) %>% 
#   summarise(n = n()) %>% 
#   mutate(prop = n/sum(n)) 
# 
# # combine the polling data proportions with the post-stratification data
# poll_post_biden <- inner_join(polling_props, voting_biden, by = "stateicp")
# poll_post_biden <- poll_post_biden %>% 
#   filter(vote_biden == 1)
# 
# # plot for predictions
# 
# poll_post_biden %>% 
#   ggplot(aes(x = employ, y = biden_predict)) +
#   geom_point(aes(color = "black")) +
#   geom_line(aes(group = 1)) +
#   scale_y_continuous(labels = scales::percent) +
#   theme(axis.text.x = element_text(angle = 70, size = 6, hjust = 1)) +
#   geom_ribbon(aes(group = 1, ymin=biden_predict_lower, ymax=biden_predict_upper), 
#               alpha = 0.2, fill = "red") +
#   geom_line(data = poll_post_biden, aes(x=stateicp, y=prop, group = 2), 
#             linetype = "twodash") +
#   geom_point(data = poll_post_biden, aes(x=stateicp, y=prop, colour = "red")) +
#   geom_hline(aes(yintercept = 0.5), alpha = 0.5) +
#   scale_color_discrete(name = "Data", labels = c("Population", "Sample")) +
#   labs(title = "Proportion of People voting for Joe Biden by State",
#        x = "State", y = "Percentage Voting Biden")
```

@@

# # Marginal Effect 
# political_preferences_predictions <- predictions(political_preferences1) |>
#   as_tibble()
# 
# political_preferences_predictions

@@

```{r}
# library(margins)
# 
# # Marginal effects for each predictor variable
# marginal_effects <- margins(political_preferences1)
# 
# # Plotting marginal effects for race
# race_effects <- dyplot(marginal_effects, variables = "race")
# race_effects
# 
# # Plotting marginal effects for region
# region_effects <- dyplot(marginal_effects, variables = "region")
# region_effects
# 
# # Plotting marginal effects for employ
# employ_effects <- dyplot(marginal_effects, variables = "employ")
# employ_effects

@@

```{r}
# # Load required libraries
# library(bayestestR)
# 
# # Extract posterior draws from the model object
# posterior_draws <- as.matrix(political_preferences1)
# 
# # Define the predictor variables for which you want to compute marginal effects
# predictor_variables <- c("Black", "Hispanic", "Middle Eastern", "Native American", 
#                          "Other race", "Two or more races", "White", "Northeast", 
#                          "South", "West", "Homemaker", "Other employment", 
#                          "Part-time", "Permanently disabled", "Retired", 
#                          "Student", "Temporarily laid off", "Unemployed")
# 
# # Compute marginal effects
# marginal_effects <- marginal_effects(posterior_draws, variables = predictor_variables)
# 
# # Summary of marginal effects
# summary(marginal_effects)

@@

```{r}
# library(tidybayes)
# 
# # Extract draws from the Bayesian model
# draws <- spread_draws(political_preferences1)
# 
# # Plot the posterior distribution of a predictor variable
# ggplot(draws, aes(x = raceBlack)) +
#   geom_density(fill = "skyblue", color = "black") +
#   labs(x = "Black", y = "Density", title = "Posterior Distribution of Black Variable")

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-modelresults2
#| fig-cap: "Credible intervals for predictors of support for Biden 2"
# Create the model plot
model_plot <- modelplot(political_preferences1, conf_level = 0.9)

@@

## Regional Variations

@@

## COVID and Mail-in Ballots

Other than race, region, employment, and other usual socioeconomical factors, the special situation of COVID-19 and mail-in-ballots also heavily contributed to the election turnout, leading to Biden's victory.  Many Republican's shift from Trump to Biden is attributed to Trump's poor COVID responses and policies.

Mail-in-ballots also spanned many controversy over fraud election as Trump claims

>>>>

library(tidyverse)
library(here)
library(rstanarm)
library(modelsummary)
library(ggplot2)
library(knitr)
library(marginaleffects)
library(plotly)
library(tibble)
library(margins)
library(kableExtra)
```

@@

This report consists of four sections, not including the introduction. In @sec-data, we look at the data used for our report and include some tables and graphs to illustrate the different groups of people in our data. In @sec-model, we build the model and discuss its justification and explanation. The results of our predictions is highlighted in @sec-result using tables and graphs. Lastly, discussions are conducted based on the findings, which addresses the voting prediction results based on race, region, and employment status.

@@

## Data Source
In this report, we use the 2020 Cooperative Election Study (CES) (@citeCes2020) as out primary dataset. This is a long-established annual political survey of US. CES contains information about how Americans view Congress and hold their representatives accountable in elections, how they vote and their electoral experiences, and how their behaviors and experiences vary with political geography and social context. In 2020, there were 61,000 American adults completed the survey.

@@

## Features
The original CES 2020 dataset, which shows in @tbl-raw, contains 61000 observations and 5 variables, which are "votereg", "CC20_410", "race", "region", "employ".

1. votereg: whether the respondent is registered to vote.

2. CC20_410: the person that respondent vote for President of the United States.

3. race: the census region where the respondent lives.

4. region: the racial or ethnic group of the respondent.

5. employ: the current employment status of the respondent.

# read in raw data
raw_ces2020 <-  read_csv(here::here("data/raw_data/ces2020.csv"))
```

```{r}
#| label: tbl-raw
#| tbl-cap: "Preview of the raw 2020 US Cooperative Election Study (CES) dataset"
#| message: false
#| echo: false

raw_ces2020[1:5, 1:5] |>
  kable()
```

## Methodology
Since it is difficult to observe through 61000 rows with 5 variables, this report will only explore and analyze through specific aspects. The original dataset contains respondent's voter registration status and the person that respondent vote for President of the United States. This paper will only analyze respondents who are registered to vote and focus on only two voting options, Biden and Trump.

The dataset is cleaned by rename the column names, specify the class of the columns, and change the numbers in the table to the corresponding description in the codebook to improve the readability. In addition, the cleaned data removes NA employment status since there are only 34 rows, which is a small amount of missing value compared with 61000 rows. The cleaned data will be analyzed and performed using R (@citeR) with `tidyverse` (@citeTidyverse), `here` (@citeHere), `rstanarm` (@citeRstanarm), `modelsummary` (@citeModleSummary), `ggplot2` (@citeGgplot2), `knitr` (@citeKnitr), `marginaleffects` (@citeMarginalEffects), `plotly` (@citePlotly), `tibble` (@citeTibble), `margins` (@citeMargins), and `kableExtra` (@citeKableExtra). @tbl-clean-data shows a preview of the cleaned dataset. @tbl-summary is a summary of the cleaned data, showing detailed statistics about the dataset.

# read in cleaned data
ces2020 <-  read_csv(here::here("data/analysis_data/cleaned_ces2020.csv"))
```

# change column type to factor
ces2020 <-
  ces2020 |>
  mutate(
    voted_for = as_factor(voted_for),
    race = factor(
      race,
      levels = c(
        "White",
        "Black",
        "Hispanic",
        "Asian",
        "Native American",
        "Middle Eastern",
        "Two or more races",
        "Other"
      )
    ),
    region = factor(
      region,
      levels = c(
        "Northeast",
        "Midwest",
        "South",
        "West"
      )
    ),
    employ = factor(
      employ,
      levels = c(
        "Full-time",
        "Part-time",
        "Temporarily laid off",
        "Unemployed",
        "Retired",
        "Permanently disabled",
        "Homemaker",
        "Student",
        "Other"
      )
    )
  ) |>
  select(voted_for, race, region, employ)
```

```{r}
#| label: tbl-clean-data
#| tbl-cap: "Preview of the cleaned 2020 CES dataset"
#| message: false
#| echo: false

ces2020[1:5, 1:4] |>
  kable()
```

```{r}
#| label: tbl-summary
#| tbl-cap: "Statistics summary of the cleaned 2020 CES dataset"
#| message: false
#| echo: false

ces2020 |>
  summary()
```

# Model {#sec-model}

@@

-------------------------------needs some explanation
```{r}
# We could use posterior predictive checks, introduced in Section 12.4, to show that the logistic regression approach is a good choice for this circumstance.

@@

# Results {#sec-result}

@@

# generate predictions for each of our grouped cells
props$estimate <- predict.glm(political_preferences1, newdata = props,
                          type = "response")

# create a new column with number of voters from each cell voting Biden
props <- props %>% mutate(num_voters = n*estimate)

# get errors for our prediction
errors <- predict.glm(model, newdata = props,
                      type = "response", se.fit = T)
lower <- errors$fit - errors$se.fit
upper <- errors$fit + errors$se.fit

# combine the proportions and the errors we generated
props_error <- cbind(props, lower, upper)
props_error <- props_error %>% 
  rename("lower" = ...11,
         "upper" = ...12)

# create a new column with number of voters from each cell voting Biden
props_error <- props_error %>% mutate(num_voters_lower = n*lower,
                                num_voters_upper = n*upper)

# create a new dataset which contains upper and lower estimates for proportions
# of votes for each state
voting_biden <- props_error %>% mutate(biden_predict_prop=estimate*prop,
                                       biden_predict_prop_lower = lower*prop,
                                       biden_predict_prop_upper = upper*prop) %>%
  group_by(stateicp) %>%
  summarise(biden_predict = sum(biden_predict_prop),
            biden_predict_lower = sum(biden_predict_prop_lower),
            biden_predict_upper = sum(biden_predict_prop_upper))
```

```{r stateplot, echo=F, message=F, warning=F, fig.cap="Proportion of each State Voting for Biden"}
# get state by state proportions for our polling data
polling_props <- polling %>% 
  group_by(stateicp, vote_biden) %>% 
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))

# combine the polling data proportions with the post-stratification data
poll_post_biden <- inner_join(polling_props, voting_biden, by = "stateicp")
poll_post_biden <- poll_post_biden %>% 
  filter(vote_biden == 1)

# plot for predictions

poll_post_biden %>% 
  ggplot(aes(x = employ, y = biden_predict)) +
  geom_point(aes(color = "black")) +
  geom_line(aes(group = 1)) +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 70, size = 6, hjust = 1)) +
  geom_ribbon(aes(group = 1, ymin=biden_predict_lower, ymax=biden_predict_upper), 
              alpha = 0.2, fill = "red") +
  geom_line(data = poll_post_biden, aes(x=stateicp, y=prop, group = 2), 
            linetype = "twodash") +
  geom_point(data = poll_post_biden, aes(x=stateicp, y=prop, colour = "red")) +
  geom_hline(aes(yintercept = 0.5), alpha = 0.5) +
  scale_color_discrete(name = "Data", labels = c("Population", "Sample")) +
  labs(title = "Proportion of People voting for Joe Biden by State",
       x = "State", y = "Percentage Voting Biden")
```

@@

# Marginal Effect 
political_preferences_predictions <- predictions(political_preferences1) |>
  as_tibble()

political_preferences_predictions

@@

```{r}
library(margins)

# Marginal effects for each predictor variable
marginal_effects <- margins(political_preferences1)

# Plotting marginal effects for race
race_effects <- dyplot(marginal_effects, variables = "race")
race_effects

# Plotting marginal effects for region
region_effects <- dyplot(marginal_effects, variables = "region")
region_effects

# Plotting marginal effects for employ
employ_effects <- dyplot(marginal_effects, variables = "employ")
employ_effects

@@

```{r}
# Load required libraries
library(bayestestR)

# Extract posterior draws from the model object
posterior_draws <- as.matrix(political_preferences1)

# Define the predictor variables for which you want to compute marginal effects
predictor_variables <- c("Black", "Hispanic", "Middle Eastern", "Native American", 
                         "Other race", "Two or more races", "White", "Northeast", 
                         "South", "West", "Homemaker", "Other employment", 
                         "Part-time", "Permanently disabled", "Retired", 
                         "Student", "Temporarily laid off", "Unemployed")

# Compute marginal effects
marginal_effects <- marginal_effects(posterior_draws, variables = predictor_variables)

# Summary of marginal effects
summary(marginal_effects)

@@

```{r}
library(tidybayes)

# Extract draws from the Bayesian model
draws <- spread_draws(political_preferences1)

# Plot the posterior distribution of a predictor variable
ggplot(draws, aes(x = raceBlack)) +
  geom_density(fill = "skyblue", color = "black") +
  labs(x = "Black", y = "Density", title = "Posterior Distribution of Black Variable")

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-modelresults2
#| fig-cap: "Credible intervals for predictors of support for Biden 1"
# Create the model plot
model_plot <- modelplot(political_preferences1, conf_level = 0.9)

@@

## Second discussion point

@@

## Talk about mail-in ballots AND Trump's later reaction to losing
